A Data-Driven Framework for Neural Field Modeling

### 2011-01-03 ltx/first_order_framework/LaTeX_Manuscript.tex
small edit the correlation analysis
added to correlation analysis derivation

 \newcommand{\parham}[1]{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}}

@@

	R_{y_{t},y_{t+1}}(\boldsymbol{\tau}) &= \mathbf{E}\left[\mathcal{F}^{-1} \left\{e^{i\boldsymbol{\nu}^{\top}\boldsymbol{\tau}} M(\boldsymbol{\nu}) V(\boldsymbol{\nu}) \ast \left(\xi M(\boldsymbol{\nu})V(\boldsymbol{\nu}) + T_s M(\boldsymbol{\nu}) W(\boldsymbol{\nu}) \varsigma V(\boldsymbol{\nu}) \right) \right\} \right] \\

@@

	&= (R_{y_t,y_t} - \delta(\boldsymbol{\tau})\sigma_{\varepsilon}^2) \ast \mathbf{E}\left[\left(\xi  + T_s \varsigma w(\mathbf{r})\right) \right]

@@

@@

Alternative, we can write
\begin{align}
	R_{y_{t},y_{t}}(\boldsymbol{\tau}) &= \mathbf{E}\left[y_{t}\left(\mathbf{r}_n+\boldsymbol{\tau}\right)y_{t}\left(\mathbf{r}_n\right)\right]\\
	&= \mathbf{E}\left[m(\mathbf{r}_n+\boldsymbol{\tau})\ast v(\mathbf{r}) \times m(\mathbf{r}_n)\ast v(\mathbf{r})\right] + \delta(\boldsymbol{\tau})\sigma_{\varepsilon}^2 \\
	&= \mathbf{E}\left[m(\mathbf{r}_n+\boldsymbol{\tau}) \ast m(\mathbf{r}_n) \ast v^2(\mathbf{r})\right] + \delta(\boldsymbol{\tau})\sigma_{\varepsilon}^2
\end{align}
To show how we get this result, we first define
\begin{align}
	K(\boldsymbol{\nu}) &= \mathcal{F}\left\{m(\mathbf{r}_n+\boldsymbol{\tau}) \ast v(\mathbf{r}) \times m(\mathbf{r}_n) \ast v(\mathbf{r})\right\} \\
	&= \exp(i\boldsymbol{\nu}^{\top}\boldsymbol{\tau})M(\boldsymbol{\nu})V(\boldsymbol{\nu}) \ast M(\boldsymbol{\nu})V(\boldsymbol{\nu}) \\
	&= \exp(i\boldsymbol{\nu}^{\top}\boldsymbol{\tau})M(\boldsymbol{\nu})V(\boldsymbol{\nu}) \ast M(\boldsymbol{\nu})V(\boldsymbol{\nu}) \\
	&= \exp(i\boldsymbol{\nu}^{\top}\boldsymbol{\tau})M(\boldsymbol{\nu})M(\boldsymbol{\nu})V(\boldsymbol{\nu}) \ast V(\boldsymbol{\nu}).
\end{align}
Now taking the inverse Fourier transform we get
\begin{align}
	k(\mathbf{r}) = m(\mathbf{r}_n+\boldsymbol{\tau}) \ast m(\mathbf{r}_n) \ast v^2(\mathbf{r}).
\end{align}

@@

\section{Derivation of Sampling Theorem}\label{App:SamplingTheorem}
\dean{This appendix provides a derivation for equations~\ref{eq:MinimumSensorDistance} and \ref{eq:BasisFunctionSeparation}. The derivation is adapted from \citet{Sanner1992} and \citet{Scerri2009}. Consider the cortical sheet defined on the surface $\Omega$ with sensors placed on a regular grid at positions $\mathbf{r}_n$ (where $n=0,...,n_y-1$ indexes the sensors) that are separated by the distance $\Delta_y$. Noise-free observations can be considered samples from a continuous field that is filtered (convolved) with the observation kernel described by
\begin{equation}
	y(\mathbf{r}_n)=\delta(\mathbf r - \mathbf{r}_n ) v_m(\mathbf{r}),
\end{equation} 
\parham{\begin{equation}
	\mathbf y=\sum_{n=0}^{n_y-1}\delta(\mathbf r - \mathbf{r}_n ) v_m(\mathbf{r}),
\end{equation}}
where $v_m(\mathbf{r}) = v_t(\mathbf{r})\ast m(\mathbf{r})$. The Fourier transform of \parham{$\mathbf y$~}$y(\mathbf{r}_n)$ is
\begin{align}
 Y(\boldsymbol{\nu}) &= \frac{1}{\Delta_y^2}\sum_{n=0}^{n_y-1} \delta\left(\boldsymbol{\nu}-\frac{\mathbf{n}_n}{\Delta_y}\right) \ast V_m(\boldsymbol \nu) \\
	&= \frac{1}{\Delta_y^2}V_m(\boldsymbol\nu) + \frac{1}{\Delta_y^2} \sum_{n=1}^{n_y-1} V_m\left(\boldsymbol{\nu}-\frac{\mathbf{n}_n}{\Delta_y}\right),
\end{align} 
where $\mathbf{n}$ is a $n_y \times 2$ matrix, spatially indexing the sensors (the columns are different spatial dimensions such that $\mathbf{r}_n = \Delta_y \mathbf{n}$). It should be clear that $Y(\boldsymbol{\nu})$ contains $V_m(\boldsymbol\nu)$ plus $n_y-1$ copies (aliases) of $V_m(\boldsymbol\nu)$ repeated on the regular lattice, where each copy is centered at $n\Delta_y^{-1}$. Due to the lowpass action of the observation kernel, the filtered dynamic field is band-limited with a cutoff frequency $\boldsymbol\nu_{cy}$. In order for $V_m(\boldsymbol\nu)$ to be disjointed from its aliases, the amount of shift $\Delta_y^{-1}$ must be at least $2\boldsymbol\nu_{cy}$, since each alias has a spectral range of
\begin{equation}
	\boldsymbol\nu = \left[\frac{n}{\Delta_y}-\boldsymbol{\nu}_{cy},\frac{n}{\Delta_y}+\boldsymbol{\nu}_{cy}\right].
\end{equation}
Therefore, to avoid aliasing the distance between observations must satisfy
\begin{equation}\label{eq:SamplingCriteria}
	\frac{1}{\Delta_y} \ge 2\boldsymbol{\nu}_{cy}.
\end{equation}
Since the observation kernel is not an ideal filter, an oversampling parameter should be included to further separate the aliases. Rearranging equation~\ref{eq:SamplingCriteria} and adding the oversampling parameter, $\rho$, gives
\begin{equation}\label{ap:Shannon}
	\Delta_y\le\frac{1}{2\rho\boldsymbol{\nu}_{cy}},
\end{equation}
where $\rho \in \mathbb{R} \ge 1$.}

>>>>

\newcommand{\parham}[1]{\textcolor{blue}{#1}}

@@

	R_{y_{t},y_{t+1}}(\boldsymbol{\tau}) &\approx \mathbf{E}\left[\mathcal{F}^{-1} \left\{e^{i\boldsymbol{\nu}^{\top}\boldsymbol{\tau}} M(\boldsymbol{\nu}) V(\boldsymbol{\nu}) \ast \left(\xi M(\boldsymbol{\nu})V(\boldsymbol{\nu}) + T_s M(\boldsymbol{\nu}) W(\boldsymbol{\nu}) \varsigma V(\boldsymbol{\nu}) \right) \right\} \right] \\

@@

	&= \left(R_{y_t,y_t}(\boldsymbol{\tau}) - \delta(\boldsymbol{\tau}\right)\sigma_{\varepsilon}^2) \ast \left(\xi  + T_s \varsigma w(\mathbf{r})\right) \label{eq:xcorr_with_autocorr}

@@

where $R_{y_{t},y_{t}}(\boldsymbol{\tau})$ is the auto-correlation of the observations and $\delta(\cdot)$ is the Kronecker delta function.

To show how we get this result, we define the auto-correlation of the observations as
\begin{align}
	R_{y_{t},y_{t}}(\boldsymbol{\tau}) &= \mathbf{E}\left[y_{t}\left(\mathbf{r}+\boldsymbol{\tau}\right)y_{t}\left(\mathbf{r}\right)\right]\\
	&= \mathbf{E}\left[m(\mathbf{r} + \boldsymbol{\tau}) \ast v(\mathbf{r}) \times m(\mathbf{r}) \ast v(\mathbf{r})\right] + \delta(\boldsymbol{\tau})\sigma_{\varepsilon}^2 \\
	&= \mathbf{E}\left[\mathcal{F}^{-1}\left\{ \exp(i\boldsymbol{\nu}^{\top}\boldsymbol{\tau})M(\boldsymbol{\nu})V(\boldsymbol{\nu}) \ast M(\boldsymbol{\nu})V(\boldsymbol{\nu}) \right\}\right] + \delta(\boldsymbol{\tau})\sigma_{\varepsilon}^2 \\
	&= \mathbf{E}\left[\mathcal{F}^{-1}\left\{ \exp(i\boldsymbol{\nu}^{\top}\boldsymbol{\tau})M(\boldsymbol{\nu})M(\boldsymbol{\nu})V(\boldsymbol{\nu}) \ast V(\boldsymbol{\nu}) \right\}\right] + \delta(\boldsymbol{\tau})\sigma_{\varepsilon}^2 \\
	&= \mathbf{E}\left[m(\mathbf{r}+\boldsymbol{\tau}) \ast m(\mathbf{r}) \ast v^2(\mathbf{r})\right] + \delta(\boldsymbol{\tau})\sigma_{\varepsilon}^2
\end{align}

Now to isolate the kernel, the Fourier transform is taken
	\mathcal{F}\left\{R_{y_{t},y_{t+1}}(\boldsymbol{\tau})\right\} &\approx \mathcal{F}\left\{ \left(R_{y_t,y_t}(\boldsymbol{\tau}) - \delta\left(\boldsymbol{\tau}\right)\sigma_{\varepsilon}^2\right) \ast \left(\xi  + T_s \varsigma w(\mathbf{r})\right) \right\} \\
	&= \mathcal{F}\left\{ R_{y_t,y_t}(\boldsymbol{\tau}) - \delta\left(\boldsymbol{\tau}\right)\sigma_{\varepsilon}^2 \right\} \mathcal{F}\left\{ \xi  + T_s \varsigma w(\mathbf{r}) \right\}.
Now rearranging
	\frac{\mathcal{F}\left\{R_{y_{t},y_{t+1}}(\boldsymbol{\tau})\right\}}{\mathcal{F}\left\{ R_{y_t,y_t}(\boldsymbol{\tau}) - \delta\left(\boldsymbol{\tau}\right)\sigma_{\varepsilon}^2 \right\}} \approx \mathcal{F}\left\{ \xi  + T_s \varsigma w(\mathbf{r}) \right\}.
Take inverse Fourier transform and rearrange to get an approximation of the kernel
	\mathcal{F}^{-1}\left\{\frac{\mathcal{F}\left\{R_{y_{t},y_{t+1}}(\boldsymbol{\tau})\right\}}{\mathcal{F}\left\{ R_{y_t,y_t}(\boldsymbol{\tau}) - \delta\left(\boldsymbol{\tau}\right)\sigma_{\varepsilon}^2 \right\}}\right\} &\approx \xi  + T_s \varsigma w(\mathbf{r}) % \\
	 % 	(\varsigma T_s)^{-1} \mathcal{F}^{-1}\left\{\frac{\mathcal{F}\left\{R_{y_{t},y_{t+1}}(\boldsymbol{\tau})\right\}} {\mathcal{F}\left\{ R_{y_t,y_t}(\boldsymbol{\tau}) - \delta\left(\boldsymbol{\tau}\right)\sigma_{\varepsilon}^2 \right\}}\right\} - (\varsigma T_s)^{-1}\xi &\approx w(\mathbf{r})
This equation states that the support of the kernel can be estimated given knowledge of $\sigma_\varepsilon^2$. Furthermore, is we assume a local coupling (at $\boldsymbol\tau=0$) regime, either excitation or inhibition, than we don't even need to know $\sigma_\varepsilon^2$. The peak in the kernel at $\mathbf{r}=0$ will be bigger or smaller if we guess incorrectly and the kernel parameters can do the rest!

@@

@@

\section{Derivation of Sampling Theorem}\label{App:SamplingTheorem}
\dean{This appendix provides a derivation for equations~\ref{eq:MinimumSensorDistance} and \ref{eq:BasisFunctionSeparation}. The derivation is adapted from \citet{Sanner1992} and \citet{Scerri2009}. Consider the cortical sheet defined on the surface $\Omega$ with sensors placed on a regular grid at positions $\mathbf{r}_n$ (where $n=0,...,n_y-1$ indexes the sensors) that are separated by the distance $\Delta_y$. Noise-free observations can be considered samples from a continuous field that is filtered (convolved) with the observation kernel described by
\begin{equation}
	y(\mathbf{r}_n)=\delta(\mathbf r - \mathbf{r}_n ) v_m(\mathbf{r}),
\end{equation} 
where $v_m(\mathbf{r}) = v_t(\mathbf{r})\ast m(\mathbf{r})$. The observations in matrix form is described by
\parham{\begin{equation}
	\mathbf y=\sum_{n=0}^{n_y-1}\delta(\mathbf r - \mathbf{r}_n ) v_m(\mathbf{r}),
\end{equation}}
The Fourier transform of $\mathbf{y}$ is
\begin{align}
 Y(\boldsymbol{\nu}) &= \frac{1}{\Delta_y^2}\sum_{n=0}^{n_y-1} \delta\left(\boldsymbol{\nu}-\frac{\mathbf{n}_n}{\Delta_y}\right) \ast V_m(\boldsymbol \nu) \\
	&= \frac{1}{\Delta_y^2}V_m(\boldsymbol\nu) + \frac{1}{\Delta_y^2} \sum_{n=1}^{n_y-1} V_m\left(\boldsymbol{\nu}-\frac{\mathbf{n}_n}{\Delta_y}\right),
\end{align} 
where $\mathbf{n}$ is a $n_y \times 2$ matrix, spatially indexing the sensors (the columns are different spatial dimensions such that $\mathbf{r}_n = \Delta_y \mathbf{n}$). It should be clear that $Y(\boldsymbol{\nu})$ contains $V_m(\boldsymbol\nu)$ plus $n_y-1$ copies (aliases) of $V_m(\boldsymbol\nu)$ repeated on the regular lattice, where each copy is centered at $n\Delta_y^{-1}$. Due to the lowpass action of the observation kernel, the filtered dynamic field is band-limited with a cutoff frequency $\boldsymbol\nu_{cy}$. In order for $V_m(\boldsymbol\nu)$ to be disjointed from its aliases, the amount of shift $\Delta_y^{-1}$ must be at least $2\boldsymbol\nu_{cy}$, since each alias has a spectral range of
\begin{equation}
	\boldsymbol\nu = \left[\frac{n}{\Delta_y}-\boldsymbol{\nu}_{cy},\frac{n}{\Delta_y}+\boldsymbol{\nu}_{cy}\right].
\end{equation}
Therefore, to avoid aliasing the distance between observations must satisfy
\begin{equation}\label{eq:SamplingCriteria}
	\frac{1}{\Delta_y} \ge 2\boldsymbol{\nu}_{cy}.
\end{equation}
Since the observation kernel is not an ideal filter, an oversampling parameter should be included to further separate the aliases. Rearranging equation~\ref{eq:SamplingCriteria} and adding the oversampling parameter, $\rho$, gives
\begin{equation}\label{ap:Shannon}
	\Delta_y\le\frac{1}{2\rho\boldsymbol{\nu}_{cy}},
\end{equation}
where $\rho \in \mathbb{R} \ge 1$.}
